---
title: "martorell 2021"
author: "Nicolas Ayala"
date: "10-01-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

##Load dataset (columns and rows as a rectangule of data)
```{r}
#Load dataframe
library(readxl)
df_martorell_2021 <- read_excel("DADES_TD_AYALA_051221.xlsx", range = "A4:G277", col_types = c("date", "numeric", "numeric", 
        "numeric", "numeric", "numeric", 
        "numeric"))
View(df_martorell_2021)

#Rename cols (7 cols)
colnames(df_martorell_2021)<-c("date","cases_attended","total_hospitalized", "patientes_pending_pcr","new_cases_user", "new_cases_professional", "patient_discharge")

#Add col of consecutive day
library(tibble)
df_martorell_2021= add_column(df_martorell_2021, consecutive_days = 1:273, .after = "date")

#Add total cases 
library(dplyr)
df_martorell_2021 <- df_martorell_2021 %>% rowwise() %>%
  mutate(total_newcases_day = sum(c_across("new_cases_user":"new_cases_professional")))

#Moving column("patient_discharge") at the end. In this way, "total_newcases_day" is together to"patientes_pending_pcr" and "new_cases_user"
df_martorell_2021<-df_martorell_2021 %>% relocate(patient_discharge, .after = patientes_pending_pcr)

```

##Variables
```{r}
##Showing the cols names
names(df_martorell_2021)

##Sample Size and numbers of variables
dim(df_martorell_2021)

##Null Values: False
table(is.null(df_martorell_2021))

##Class of Cols.
sapply(df_martorell_2021, class)
```


##Summary and descriptive analysis
```{r}
## Min, 1Q, Median, Mean, 3Q and Max Values
summary(df_martorell_2021)

## Summary of Variables
library(Hmisc)
describe(df_martorell_2021)

```
##Distribution of data
```{r}
## Distribution-Hospital of Martorell 20221 (january-agost)
layout(matrix(c(1:5), nrow=1))
boxplot(df_martorell_2021$total_newcases_day, col="lightgreen", main="Daily New Cases", ylab= "N°Cases", ylim= c(0,25))
boxplot(df_martorell_2021$cases_attended, col="lightgreen", main= "Total attended", ylim= c(0,25))
boxplot(df_martorell_2021$total_hospitalized, col="lightgreen", main="Daily Hospitalized",  ylim= c(0,25))
boxplot(df_martorell_2021$patient_discharge, col="lightgreen", main="Daily Discharge", ylim= c(0,25))
boxplot(df_martorell_2021$patientes_pending_pcr, col="lightgreen", main="PCR in process", ylim= c(0,25))
```
##Plot of Variables
```{r}
#1.Cases per consecutive days
library(ggplot2)
plot1<-ggplot(data=df_martorell_2021, aes(y=total_newcases_day, x=date)) +
  geom_line(color="steelblue")+
   theme_bw()+
   ylim(0, 15)+
   ggtitle (" Daily new cases.")+
   xlab("2021")+
   ylab("Cases")


#2.Cases in Hospital of Martorell
library(ggplot2)
plot2<-ggplot(data=df_martorell_2021, aes(y=cases_attended, x=date)) +
  geom_line(color="steelblue")+
   theme_bw()+
   ylim(0, 15)+
   ggtitle (" Total Cases in all Services.")+
   xlab("2021")+
   ylab("Cases")


#Cases COVID-19 Hospitalized
library(ggplot2)
plot3<-ggplot(data=df_martorell_2021, aes(y=total_hospitalized, x=date)) +
  geom_line(color="steelblue")+
   theme_bw()+
   ylim(0, 15)+
   ggtitle (" COVID-19 Hospitalized.")+
   xlab("2021")+
   ylab("Cases")

#Patient Discharged
library(ggplot2)
plot4<-ggplot(data=df_martorell_2021, aes(y=patient_discharge, x=date)) +
  geom_line(color="steelblue")+
   theme_bw()+
   ylim(0, 15)+
   ggtitle (" Discharged Patient.")+
   xlab("2021")+
   ylab("Cases")

plot1
plot2
plot3
plot4


```

#Cowplot for New Cases, Total Cases, Hospitalized and Discharged Patient
```{r}
library(cowplot)
plot_grid(plot1,plot2,plot3,plot4,labels = c("A", "B", "C", "D"), ncol = 2, nrow = 2)

```


## Total cases and new cases
```{r}
library(ggplot2)
ggplot(data=df_martorell_2021) +
   geom_line(aes(x=date,y=df_martorell_2021$total_newcases_day),color='red', size=0.1)+
   geom_col(aes(x=date,y=df_martorell_2021$cases_attended),fill = "#00abff", alpha=0.5)+
   ylim(0,25)+
   ylab('Cases')+xlab('Date 2021')+
  theme_bw()+
   ggtitle("Cases COVID-19 in Hospital of Martorell")

```


##Total Cases, new cases and hospitalized patients.
## Otra opcion es suavizar la linea con la media de 7 dias
```{r}
library(ggplot2)
ggplot(data=df_martorell_2021) +
  geom_line(aes(x=date,y=df_martorell_2021$total_newcases_day),color='red', size=0.2)+
  geom_col(aes(x=date,y=df_martorell_2021$cases_attended),fill = "#00abff", alpha=0.5)+
  geom_line(aes(x=date,y=df_martorell_2021$total_hospitalized),color=8, size=0.1)+
  theme_bw()+
  ylim(0,25)+
  ylab('Cases')+xlab('Date 2021')+
  ggtitle("Cases COVID-19 in Hospital of Martorell")


```
## Comparison with Martorell Data
## Web source (12/01/2021): https://dadescovid.cat/descarregues?lang=cas&tipus_territori=territori&scol=EXITUS&id_html=up_1_5&codi=08114&tipus=municipi
```{r}
#Import dataset.
library(readr)
municipi <- read_delim("casos_municipi.csv", delim = ";", escape_double = FALSE, col_types = cols(TIPUSCASDATA = col_date(format = "%d/%m/%Y")), trim_ws = TRUE)
#View(municipi)
martorell_town<-filter (municipi,MUNICIPIDESCRIPCIO  == "MARTORELL") 

##Summarise per date
library(dplyr)
martorell_town<-martorell_town %>%group_by(TIPUSCASDATA) %>%
summarise(TOTCASOS = sum(NUMCASOS))


#Rename cols (7 cols)
colnames(martorell_town)<-c("date","total_cases")

#Add col of consecutive day
library(dplyr)
martorell_town$consecutive_days <- 1:nrow(martorell_town) 
martorell_town <- martorell_town %>% relocate(consecutive_days, .before = total_cases)

#head of Dataframe
head(martorell_town)

```


##Martorell Township Variables
```{r}
##Showing the cols names
names(martorell_town)

##Sample Size and numbers of variables: 1062 rows, 7 cols
dim(martorell_town)

##Null Values: False
table(is.null(martorell_town))

##Class of Cols.
sapply(martorell_town, class)
```


```{r}
##COVID-19 cases in 2020-2022
## Personas las personas vacunadas dentro del hospital. 

library(ggplot2)
a<-ggplot(data= martorell_town , aes(date, total_cases))+
  geom_col()+
  ggtitle("Cases in Martorell: 2020-2022")+
  ylim(0,35)


b<-ggplot(data= martorell_town , aes(date, total_cases))+
  geom_col()+
  scale_x_date(limits = as.Date(c("2020-01-01", "2020-09-30")))+
  ggtitle("Cases in Martorell: Jan-Sept 2020")+
  ylim(0,35)

##COVID cases in 2021 (jan-sept)
c<-ggplot(data= martorell_town , aes(date, total_cases))+
  geom_col()+
  scale_x_date(limits = as.Date(c("2021-01-01", "2021-09-30")))+
  ggtitle("Cases in Martorell: Jan-Sept 2021")+
  ylim(0, 35)

##
library(cowplot)
plot_grid(b,c, labels = c("A", "B", "C"), ncol = 1, nrow = 3)
         
```


##Summary and descriptive analysis
```{r}
## Min, 1Q, Median, Mean, 3Q and Max Values
summary(martorell_town)

## Summary of Variables
library(Hmisc)
describe(martorell_town)

```


##Linear regression
```{r}
##A hypothesis could be that when cases increase, mortality increases. This is NOT a valid conclusion. A simple regression of deaths vs. cases and time could shows the passage of time has more explanatory power than cases in predicting deaths. We need number of deaths Plot ideas: Comparison 2020 vs 2021,Comparison Catalonia, Barcelona, Martorell.

# passage of time affects deaths more than cases (?)
lm(mortality_7days ~ cases_7day + date, data = df_martorell_2021) %>%
  tidy() %>%
  gt() %>% 
  tab_options(table.width = "60%") %>%
  tab_style(style = table_style, 
            locations = cells_body()) %>% 
  opt_all_caps()

```



##Predictive Model

##Idea: To check predective model: Linear regresion, Multivariate, Bayes or Montecarlo.

## Testing predective model covid.cases.cumcases3(BDSbiost3): Using first fifty days of new cases.
```{r}
# Predective model 
library(BDSbiost3)
pred.cases.fifty_days<-covid.cases.cumcases3(X=df_martorell_2021[1:50,]$consecutive_days, Y=df_martorell_2021[1:50,]$total_newcases_day, max.cases.free=T, max.cases = max(df_martorell_2021[1:50,]$total_newcases_day) , lim.days=60,iterations = 20,label.ind="Cases COVID-19", duration_pandemic=55,join.mod2 = F)

# Graph Cases for 50 days"
library(ggplot2)
ggplot(data= df_martorell_2021[1:60,], aes(consecutive_days, total_newcases_day))+
  geom_line(aes(colour=consecutive_days))+
  ggtitle("Cases in Martorell: 60 days")+
  ylim(0,12)+
  scale_y_continuous(breaks = seq(0, 12, by = 1))+
  scale_x_continuous(breaks = seq(0, 60, by = 10))+
  geom_vline(xintercept=which(df_martorell_2021$consecutive_days == '50'), linetype="dashed" , alpha= 0.5)


# Range between 50-60 day: All days. 
r50_60<- range(df_martorell_2021[50:60,]$total_newcases_day)
print(r50_60)

#The range obtained is within the confidence interval except one point (day 56)
```

## 100-110 days
```{r}
## Pred
library(ggplot2)
library(BDSbiost3)
pred.cases.onehundred<-covid.cases.cumcases3(X=df_martorell_2021[1:100,]$consecutive_days, Y=df_martorell_2021[1:100,]$total_newcases_day, max.cases.free=T, max.cases = max(df_martorell_2021[1:100,]$total_newcases_day) , lim.days=110,iterations = 20,label.ind="Cases COVID-19", duration_pandemic=110,join.mod2 = F)

library(ggplot2)
ggplot(data= df_martorell_2021[1:110,], aes(consecutive_days, total_newcases_day))+
geom_line(aes(colour=consecutive_days))+
ggtitle("Cases in Martorell: 110 days")+
ylim(0,12)+
scale_y_continuous(breaks = seq(0, 12, by = 1))+
scale_x_continuous(breaks = seq(0, 110, by = 10))+
geom_vline(xintercept=which(df_martorell_2021$consecutive_days == '100'), linetype="dashed" , alpha= 0.5)

r100_110<- range(df_martorell_2021[100:110,]$total_newcases_day)
print(r100_110)

#The range obtained is within the confidence interval.


```


## 200-210 days
```{r}
library(BDSbiost3)
pred.cases.onehundred<-covid.cases.cumcases3(X=df_martorell_2021[1:200,]$consecutive_days, Y=df_martorell_2021[1:200,]$total_newcases_day, max.cases.free=T, max.cases = max(df_martorell_2021[1:200,]$total_newcases_day) , lim.days=210,iterations = 20,label.ind="Cases COVID-19", duration_pandemic=110,join.mod2 = F)

library(ggplot2)
ggplot(data= df_martorell_2021[1:210,], aes(consecutive_days, total_newcases_day))+
  geom_line(aes(colour=consecutive_days))+
  ggtitle("Cases in Martorell: 200 days")+
  ylim(0,18)+
  scale_y_continuous(breaks = seq(0, 20, by = 1))+
  scale_x_continuous(breaks = seq(0, 210, by = 10))+
    geom_vline(xintercept=which(df_martorell_2021$consecutive_days == '200'), linetype="dashed" , alpha= 0.5)

r200_210<- range(df_martorell_2021[200:110,]$total_newcases_day)
print(r200_210)

#The range obtained is not within the confidence interval.


```

## 250-260 days
```{r}
library(BDSbiost3)
pred.cases.onehundred<-covid.cases.cumcases3(X=df_martorell_2021[1:250,]$consecutive_days, Y=df_martorell_2021[1:250,]$total_newcases_day, max.cases.free=T, max.cases = max(df_martorell_2021[1:250,]$total_newcases_day) , lim.days=260,iterations = 20,label.ind="Cases COVID-19", duration_pandemic=110,join.mod2 = F)

library(ggplot2)
ggplot(data= df_martorell_2021[1:260,], aes(consecutive_days, total_newcases_day))+
  geom_line(aes(colour=consecutive_days))+
  ggtitle("Cases in Martorell: 260 days")+
  ylim(0,18)+
  xlim(200,260)+
  scale_y_continuous(breaks = seq(0, 20, by = 1))+
  scale_x_continuous(breaks = seq(0, 260, by = 10))+
  geom_vline(xintercept=which(df_martorell_2021$consecutive_days == '250'), linetype="dashed" , alpha= 0.5)

r250_260<- range(df_martorell_2021[250:260,]$total_newcases_day)
print(r250_260)

#The range obtained is within the confidence interval. However, the regression line is increasing
```


```{r}

#Cases in hospital
pred.Casos.diarios<-covid.cases.cumcases3(X=df_martorell_2021$consecutive_days, Y=df_martorell_2021$cases_attended, max.cases.free=T, max.cases = max(df_martorell_2021$cases_attended) , lim.days=290,iterations = 100,label.ind="Cases COVID-19", duration_pandemic=300,join.mod2 = F)
```

## Comentarios de Script original con Toni en UB
## Ingresar datos del modelo por olas, buscar las lineas de tendencia y luego proyectar la incidencia e incluir los datos en el modelo.
## Multivariate time series model (MTS).
## Require(MTS)
## Frequenncy: Cada 24 horas en el modelo 

## Multivariate time series model (MTS).
## Reference documentation:
## https://cran.r-project.org/web/packages/MTS/index.html
## https://bookdown.org/singh_pratap_tejendra/intro_time_series_r/multivariate-ts-analysis.html

```{r}
# Packages

require(tidyverse)
require(tidymodels)
require(data.table)
require(tidyposterior)
require(tsibble)  #tsibble for time series based on tidy principles
require(fable)  #for forecasting based on tidy principles
require(ggfortify)  #for plotting timeseries
require(forecast)  #for forecast function
require(tseries)
require(chron)
require(lubridate)
require(directlabels)
require(zoo)
require(lmtest)
require(TTR)  #for smoothing the time series
require(MTS)
require(vars)
require(fUnitRoots)
require(lattice)
require(grid)

```


```{r}
# Attaching the Martorell Hospital database. Convert de data variable and deleting columns: date, consecutive_days, new_cases_user and new_cases_professional.
data<-data.frame(df_martorell_2021)
data$date <- as.Date(data$date, format = "%Y-%m-%d")
data$date <- NULL
data$consecutive_days<-NULL
data$new_cases_user<-NULL
data$new_cases_professional<-NULL

data100<- data[1:100,]
head(data100)

ts_COVID19 = ts(data100)
class(ts_COVID19)
plot(ts_COVID19)

```

```{r}
#
theme_set(theme_bw())
autoplot(ts_COVID19) +
  ggtitle("Time Series Plot of the `mymts' Time-Series") +
  theme(plot.title = element_text(hjust = 0.4)) #for centering the text


```

```{r}
# Main packages - problem: both have different functions VAR
## Testing for stationarity
### tseries - standard test adt.test
apply(ts_COVID19, 2, adf.test)

```

```{r}
# Alternative: lib fUnitRoots, function
apply(ts_COVID19, 2, adfTest, 
      lags=0, #maximum number of lags used for error term correction
      type="c", #type of unit root regression
      title = "ADF Test for COVID-19 Data") #title of the project
```

```{r}
# Differencing the whole mts
stnry = diffM(ts_COVID19) #difference operation on a vector of time series. Default order of differencing is 1.

# Retest
apply(ts_COVID19, 2, adf.test)
```

```{r}
## VAR modeling
plot.ts(stnry)

autoplot(ts(stnry)) +
  ggtitle("Time Series Plot of the stationary `COVID-19' Time-Series")
```
```{r}
# Lag order identification
#We will use two different functions, from two different packages to identify the lag order for the VAR model. Both functions are quite similar to each other but differ in the output they produce. vars::VAR is a more powerful and convinient function to identify the correct lag order. 

VARselect(stnry, 
          type = "none", #type of deterministic regressors to include. We use none because the time series was made stationary using differencing above. 
          lag.max = 10) #highest lag order
```

```{r}
# Creating a VAR model with vars
var.a <- vars::VAR(stnry,
                   lag.max = 10, #highest lag order for lag length selection according to the choosen ic
                   ic = "AIC", #information criterion
                   type = "none") #type of deterministic regressors to include
summary(var.a)
```

```{r}
# Residual diagnostics
# serial.test function takes the VAR model as the input.  
serial.test(var.a)
```

```{r}
# selecting the variables
# Granger test for causality
# for causality function to give reliable results we need all the variables of the multivariate time series to be stationary. 

causality(var.a, #VAR model
          cause = c("cases_attended")) 

#cause variable. If not specified then first column of x is used. Multiple variables can be used. 
```

```{r}
## Forecasting VAR models
fcast = predict(var.a, n.ahead = 10) # we forecast over a short horizon because beyond short horizon prediction becomes unreliable or uniform
par(mar = c(2.5,2.5,2.5,2.5))
plot(fcast)
```

```{r}
# Forecasting the Cases attended 
cases_attended = fcast$fcst[1] 
cases_attended # type list
class(cases_attended)

```


```{r}
# Extracting the forecast column
x = cases_attended$cases_attended[,1]
x
```

```{r}
tail(ts_COVID19)
# Inverting the differencing
#To get the data to the original scale we invert the time series
#since the values are just difference from the previous value, to get the values on the original scale we add the last value from the Cases Attended time series to the predicted values.
#the plot of the predicted values will also show that over longer horizon the predicted values are not reliable
x = cumsum(x)+5
x
par(mar = c(2.5,2.5,1,2.5)) #bottom, left, top, and right
plot.ts(x)
```

```{r}
# Adding data and forecast to one time series
cattended_inv =ts(c(ts_COVID19[,1], x))

plot(cattended_inv)
plot.ts(cattended_inv[100:110])

```

```{r}
cattended_inv_df <- as.data.frame(cattended_inv[1:110]) 
colnames(cattended_inv_df) <- c("x")
head(cattended_inv_df)
tail(cattended_inv_df)

```


```{r}
#Pending result in chunk
class(cattended_inv_df)

df.real.attended= as.data.frame(df_martorell_2021$cases_attended)
df.real.attended

#Verification of Forecast with real data during 101-110 days.
review_fcast.attended = cbind(cattended_inv_df[1:110, ] , df.real.attended[1:110, ])
review_fcast.attended = as.data.frame(review_fcast.attended)
names(review_fcast.attended)[1] <- "forecast"
names(review_fcast.attended)[2] <- "cases"


#Testing Error of Forecast
review_fcast.attended<-review_fcast.attended %>% 
  mutate(RPD = abs((cases-forecast)/((cases+forecast)/2)*100))

review_fcast.attended <- review_fcast.attended[101:110, ]
review_fcast.attended
mean(review_fcast.attended$RPD)
sd(review_fcast.attended$forecast)
sd(review_fcast.attended$cases)


#Ploting the forecast
library(ggplot2)
a=
ggplot() + 
  geom_line(data = as.data.frame(cattended_inv_df[1:100, ]), aes(y = get("cattended_inv_df[1:100, ]"), x = seq(1, 100)), color = "green") +
  geom_line(data = as.data.frame(cattended_inv_df[101:110, ]), aes(y = get("cattended_inv_df[101:110, ]"), x = seq(101, 110)), color = "red")+
  
  geom_line(data = as.data.frame(df.real.attended[101:110, ]), aes(y = get("df.real.attended[101:110, ]"), x = seq(101, 110)), color = "blue",linetype=2) +
  
   theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Time: Days") + ylab("Cases")+
  ggtitle("Cases Attended")
a

ggplot()+
  geom_line(data = as.data.frame(cattended_inv_df[101:110, ]), aes(y = get("cattended_inv_df[101:110, ]"), x = seq(101, 110)), color = "red")+
  
  geom_line(data = as.data.frame(df.real.attended[101:110, ]), aes(y = get("df.real.attended[101:110, ]"), x = seq(101, 110)), color = "blue",linetype=2) +
  
   theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Time") + ylab("Value")
```





## Total Hospitalized
Forecasting new cases per day.
```{r New Cases per Day}
# Forecasting new cases per day
totalhosp = fcast$fcst[2] 
totalhosp # type list
class(totalhosp)

```


```{r}
# Extracting the forecast column
x.totalhosp= totalhosp$total_hospitalized[,1]
x.totalhosp
```

```{r}
tail(ts_COVID19)
# Inverting the differencing
#To get the data to the original scale we invert the time series
#since the values are just difference from the previous value, to get the values on the original scale we add the last value from the Cases Attended time series to the predicted values.
#the plot of the predicted values will also show that over longer horizon the predicted values are not reliable
x.totalhosp = cumsum(x.totalhosp)+8
x.totalhosp
par(mar = c(2.5,2.5,1,2.5)) #bottom, left, top, and right
plot.ts(x.totalhosp)
```

```{r}
# Adding data and forecast to one time series
totalhosp_inv =ts(c(ts_COVID19[,2], x.totalhosp))

plot(totalhosp_inv)
plot.ts(totalhosp_inv[100:110])

```

```{r}
totalhosp_inv_df <- as.data.frame(totalhosp_inv[1:110]) 
colnames(totalhosp_inv_df) <- c("x")
head(totalhosp_inv_df)
tail(totalhosp_inv_df)

```


```{r}
#Forecast Total Hospitalized
class(totalhosp_inv_df)

df.total.hosp= as.data.frame(df_martorell_2021$total_hospitalized)
df.total.hosp

#Verification of Forecast with real data during 101-110 days. Point of Control
review_fcast.hosp = cbind(totalhosp_inv_df[1:110, ] , df.total.hosp[1:110, ])
review_fcast.hosp

#Verification of Forecast with real data during 101-110 days. Point of Control
review_fcast.hosp = cbind(totalhosp_inv_df[1:110, ] , df.total.hosp[1:110, ])
review_fcast.hosp= as.data.frame(review_fcast.hosp)
names(review_fcast.hosp)[1] <- "forecast"
names(review_fcast.hosp)[2] <- "cases"


#Testing Error of Forecast
review_fcast.hosp<-review_fcast.hosp %>% 
  mutate(RPD = abs((cases-forecast)/((cases+forecast)/2)*100))

review_fcast.hosp <- review_fcast.hosp[101:110, ]
review_fcast.hosp
mean(review_fcast.hosp$RPD)
sd(review_fcast.hosp$forecast)
sd(review_fcast.hosp$cases)


#Ploting the forecast
library(ggplot2)
b= ggplot() + 
  geom_line(data = as.data.frame(totalhosp_inv_df[1:100, ]), aes(y = get("totalhosp_inv_df[1:100, ]"), x = seq(1, 100)), color = "green") +
  geom_line(data = as.data.frame(totalhosp_inv_df[101:110, ]), aes(y = get("totalhosp_inv_df[101:110, ]"), x = seq(101, 110)), color = "red")+
  
  geom_line(data = as.data.frame(df.total.hosp[101:110, ]), aes(y = get("df.total.hosp[101:110, ]"), x = seq(101, 110)), color = "blue",linetype=2) +
  
   theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Time: Days") + ylab("Cases")+
  ggtitle("Hospitalized Patients")
b

ggplot()+
  geom_line(data = as.data.frame(totalhosp_inv_df[101:110, ]), aes(y = get("totalhosp_inv_df[101:110, ]"), x = seq(101, 110)), color = "red")+
  
  geom_line(data = as.data.frame(df.total.hosp[101:110, ]), aes(y = get("df.total.hosp[101:110, ]"), x = seq(101, 110)), color = "blue",linetype=2) +
  
   theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Time: Days") + ylab("Cases")
```






## pending PCR
Forecasting new cases per day.
```{r New Cases per Day}
# Forecasting pending PCR per patient
ppcr = fcast$fcst[3] 
ppcr # type list
class(ppcr)

```


```{r}
# Extracting the forecast column
x.ppcr= ppcr$patientes_pending_pcr[,1]
x.ppcr
```

```{r}
tail(ts_COVID19)
# Inverting the differencing
#To get the data to the original scale we invert the time series
#since the values are just difference from the previous value, to get the values on the original scale we add the last value from the Cases Attended time series to the predicted values.
#the plot of the predicted values will also show that over longer horizon the predicted values are not reliable
x.ppcr = cumsum(x.ppcr)+1
x.ppcr
par(mar = c(2.5,2.5,1,2.5)) #bottom, left, top, and right
plot.ts(x.ppcr)
```

```{r}
# Adding data and forecast to one time series
ppcr_inv =ts(c(ts_COVID19[,3], x.ppcr))

plot(ppcr_inv)
plot.ts(ppcr_inv[100:110])

```

```{r}
ppcr_inv_df <- as.data.frame(ppcr_inv[1:110]) 
colnames(ppcr_inv_df) <- c("x")
head(ppcr_inv_df)
tail(ppcr_inv_df)

```


```{r}
#Pending result in chunk
class(ppcr_inv_df)

df.p.pcr= as.data.frame(df_martorell_2021$patientes_pending_pcr)
df.p.pcr

#Verification of Forecast with real data during 101-110 days. Point of Control
review_fcast.ppcr = cbind(ppcr_inv_df[1:110, ] , df.p.pcr[1:110, ])
review_fcast.ppcr = as.data.frame(review_fcast.ppcr)
names(review_fcast.ppcr)[1] <- "forecast"
names(review_fcast.ppcr)[2] <- "cases"


#Testing Error of Forecast
review_fcast.ppcr<-review_fcast.ppcr %>% 
  mutate(RPD = abs((cases-forecast)/((cases+forecast)/2)*100))

review_fcast.ppcr <- review_fcast.ppcr[101:110, ]
review_fcast.ppcr
mean(review_fcast.ppcr$RPD)
sd(review_fcast.ppcr$forecast)
sd(review_fcast.ppcr$cases)

#Ploting the forecast
library(ggplot2)
c=
ggplot() + 
  geom_line(data = as.data.frame(ppcr_inv_df[1:100, ]), aes(y = get("ppcr_inv_df[1:100, ]"), x = seq(1, 100)), color = "green") +
  geom_line(data = as.data.frame(ppcr_inv_df[101:110, ]), aes(y = get("ppcr_inv_df[101:110, ]"), x = seq(101, 110)), color = "red")+
  
  geom_line(data = as.data.frame(df.p.pcr[101:110, ]), aes(y = get("df.p.pcr[101:110, ]"), x = seq(101, 110)), color = "blue",linetype=2) +
  
   theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Time: Days") + ylab("Cases")+
  ggtitle("PCR Pending")
c

ggplot()+
  geom_line(data = as.data.frame(ppcr_inv_df[101:110, ]), aes(y = get("ppcr_inv_df[101:110, ]"), x = seq(101, 110)), color = "red")+
  
  geom_line(data = as.data.frame(df.p.pcr[101:110, ]), aes(y = get("df.p.pcr[101:110, ]"), x = seq(101, 110)), color = "blue",linetype=2) +
  
   theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Time") + ylab("Value")
```




## Patient Discharged
Forecasting Patient Discharged per Day.
```{r New Cases per Day}
# Forecasting pending PCR per patient
pdischarge = fcast$fcst[4] 
pdischarge # type list
class(pdischarge)

```


```{r}
# Extracting the forecast column
x.pdischarge= pdischarge$patient_discharge[,1]
x.pdischarge
```

```{r}
tail(ts_COVID19)
# Inverting the differencing
#To get the data to the original scale we invert the time series
#since the values are just difference from the previous value, to get the values on the original scale we add the last value from the Cases Attended time series to the predicted values.
#the plot of the predicted values will also show that over longer horizon the predicted values are not reliable
x.pdischarge = cumsum(x.pdischarge)+3
x.pdischarge
par(mar = c(2.5,2.5,1,2.5)) #bottom, left, top, and right
plot.ts(x.pdischarged)
```

```{r}
# Adding data and forecast to one time series
pdischarge_inv =ts(c(ts_COVID19[,4], x.pdischarge))

plot(pdischarge_inv)
plot.ts(pdischarge_inv[100:110])

```

```{r}
pdischarge_inv_df <- as.data.frame(pdischarge_inv[1:110]) 
colnames(pdischarge_inv_df) <- c("x")
head(pdischarge_inv_df)
tail(pdischarge_inv_df)

```


```{r}
#Forecast Discharge Patient.
class(pdischarge_inv_df)

df.p.discharge= as.data.frame(df_martorell_2021$patient_discharge)
df.p.discharge

#Verification of Forecast with real data during 101-110 days. Point of Control
review_fcast.discharge = cbind(pdischarge_inv_df[1:110, ] , df.p.discharge[1:110, ])
review_fcast.discharge = as.data.frame(review_fcast.discharge)
class(review_fcast.discharge$cases)
names(review_fcast.discharge)[1] <- "forecast"
names(review_fcast.discharge)[2] <- "cases"


#Testing Error of Forecast
review_fcast.discharge<-review_fcast.discharge %>% 
  mutate(RPD = abs((cases-forecast)/((cases+forecast)/2)*100))

bias.fcast.discharge <- review_fcast.discharge[101:110, ]
bias.fcast.discharge
mean(bias.fcast.discharge$RPD)
sd(bias.fcast.discharge$forecast)
sd(bias.fcast.discharge$cases)


#Ploting the forecast
library(ggplot2)
d=
ggplot() + 
  geom_line(data = as.data.frame(pdischarge_inv_df[1:100, ]), aes(y = get("pdischarge_inv_df[1:100, ]"), x = seq(1, 100)), color = "green") +
  geom_line(data = as.data.frame(pdischarge_inv_df[101:110, ]), aes(y = get("pdischarge_inv_df[101:110, ]"), x = seq(101, 110)), color = "red")+
  
  geom_line(data = as.data.frame(df.p.discharge[101:110, ]), aes(y = get("df.p.discharge[101:110, ]"), x = seq(101, 110)), color = "blue",linetype=2) +
  
   theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Time: Days") + ylab("Cases")+
  ggtitle("Discharge Patient")
d

ggplot()+
  geom_line(data = as.data.frame(pdischarge_inv_df[101:110, ]), aes(y = get("pdischarge_inv_df[101:110, ]"), x = seq(101, 110)), color = "red")+
  
  geom_line(data = as.data.frame(df.p.discharge[101:110, ]), aes(y = get("df.p.discharge[101:110, ]"), x = seq(101, 110)), color = "blue",linetype=2) +
  
   theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Time") + ylab("Value")
```

## New Cases Per Day
Forecasting new cases per day.
```{r New Cases per Day}
# Forecasting new cases per day
newcases = fcast$fcst[5] 
newcases # type list
class(newcases)

```


```{r}
# Extracting the forecast column
x.newcases = newcases$total_newcases_day[,1]
x.newcases
```

```{r}
tail(ts_COVID19)
# Inverting the differencing
#To get the data to the original scale we invert the time series
#since the values are just difference from the previous value, to get the values on the original scale we add the last value from the Cases Attended time series to the predicted values.
#the plot of the predicted values will also show that over longer horizon the predicted values are not reliable
x.newcases = cumsum(x.newcases)+2
x.newcases
par(mar = c(2.5,2.5,1,2.5)) #bottom, left, top, and right
plot.ts(x.newcases)
```

```{r}
# Adding data and forecast to one time series
newcases_inv =ts(c(ts_COVID19[,5], x.newcases))

plot(newcases_inv)
plot.ts(newcases_inv[100:110])

```

```{r}
newcases_inv_df <- as.data.frame(newcases_inv[1:110]) 
colnames(newcases_inv_df) <- c("x")
head(newcases_inv_df)
tail(newcases_inv_df)

```


```{r}
#Pending result in chunk
class(newcases_inv_df)

df.new.cases= as.data.frame(df_martorell_2021$total_newcases_day)
df.new.cases

#Verification of Forecast with real data during 101-110 days.
review_fcast.ncases = cbind(newcases_inv_df[1:110, ] , df.new.cases[1:110, ])
review_fcast.ncases= as.data.frame(review_fcast.ncases)

names(review_fcast.ncases)[1] <- "forecast"
names(review_fcast.ncases)[2] <- "cases"
review_fcast.ncases<- review_fcast.ncases[101:110, ]

review_fcast.ncases<-review_fcast.ncases %>% 
  mutate(RPD = abs((cases-forecast)/((cases+forecast)/2)*100))


review_fcast.ncases
mean(review_fcast.ncases$RPD)
sd(review_fcast.ncases$forecast)
sd(review_fcast.ncases$cases)

```


```{r}

#Ploting the forecast
library(ggplot2)
e=
ggplot() + 
  geom_line(data = as.data.frame(newcases_inv_df[1:100, ]), aes(y = get("newcases_inv_df[1:100, ]"), x = seq(1, 100)), color = "green") +
  geom_line(data = as.data.frame(newcases_inv_df[101:110, ]), aes(y = get("newcases_inv_df[101:110, ]"), x = seq(101, 110)), color = "red")+
  
  geom_line(data = as.data.frame(df.new.cases[101:110, ]), aes(y = get("df.new.cases[101:110, ]"), x = seq(101, 110)), color = "blue",linetype=2) +
  
   theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Time: Days") + ylab("Cases")+
  ggtitle("New Cases")
e

ggplot()+
  geom_line(data = as.data.frame(newcases_inv_df[101:110, ]), aes(y = get("newcases_inv_df[101:110, ]"), x = seq(101, 110)), color = "red")+
  
  geom_line(data = as.data.frame(df.new.cases[101:110, ]), aes(y = get("df.new.cases[101:110, ]"), x = seq(101, 110)), color = "blue",linetype=2) +
  
   theme(plot.title = element_text(hjust = 0.5)) +
  xlab("Time") + ylab("Cases")


```

```{r}
install.packages("ggpubr")
library(ggpubr)

ggarrange(a,b, c, d, e , 
          labels = c("A", "B", "C", "D", "E"),
          ncol = 2, nrow = 3)
```


